
```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#knitr::opts_chunk$set(echo = TRUE)
```

### Trend of US Immigration Enforcement

In the coming month I'll be digging into the immigration enforcement data posted on [data.world](https://data.world/stoltzmaniac/insights-from-the-usa). I encourage anyone to take this data and either add to the project or to do something on their own. I will be bringing in external data sources to merge as well (which I did for this first plot).

Anyway, if you're only here for a "high-level nugget" of information, the basic thing you can see is:  

#### Things have changed since 1925!  

Here's a look at the immigration enforcement data with the party of the POTUS shown in red and blue:  

----

```{r execute script}
library(tidyverse)
library(leaflet)
library(rgdal)
## Data saved on data.world https://data.world/stoltzmaniac/insights-from-the-usa/
#wall = geojsonio::geojson_read('../Mexico Border Wall/border-fence.geojson')
#wall = readOGR('../Mexico Border Wall/border-fence.geojson')
```


```{r,fig.width=3}
#leaflet() %>% 
#  addTiles() %>% 
#  setView(-106, 31, zoom = 5) %>%
#  addGeoJSON(wall)
```

