
```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#knitr::opts_chunk$set(echo = TRUE)
```

### Trend of US Immigration Enforcement

In the coming month I'll be digging into the immigration enforcement data posted on [data.world](https://data.world/stoltzmaniac/insights-from-the-usa). I encourage anyone to take this data and either add to the project or to do something on their own. I will be bringing in external data sources to merge as well (which I did for this first plot).

Anyway, if you're only here for a "high-level nugget" of information, the basic thing you can see is:  

#### Things have changed since 1925!  

Here's a look at the immigration enforcement data with the party of the POTUS shown in red and blue:  

----

```{r execute script}
library(data.world)
library(reshape2)
library(tidyverse)
library(gridExtra)
library(scales)
library(GGally)
library(lubridate)
library(ggthemes)

# Datasets are referenced by their URL or path
dataset_key <- "https://data.world/stoltzmaniac/insights-from-the-usa"

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", "immigrationenforcement"))
data = data.world::query(qry, dataset = dataset_key)

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", '02_2_fbi_hate_crime_statistics_comparison_2000_2015_all'))
hate_crimes = data.world::query(qry, dataset = dataset_key)

df = hate_crimes %>%
  left_join(data, by = 'year') %>%
  arrange(year)


```

```{r}
df
```



```{r}
pct.change <- function(x) {(x-lag(x))/lag(x)}
df.changes = data %>% arrange(year)
df.changes$year = as.character(df.changes$year)

df.changes = df.changes %>%
  mutate_if(is.numeric, pct.change) %>%
  na.omit() %>% 
  gather(Enforcement,`Annual % Change`,apprehended:returns) %>%
  arrange(year)

df.changes$year = as.integer(df.changes$year)
#df.changes$decade = (df.changes$year %/% 10) * 10
df.changes$Enforcement = factor(df.changes$Enforcement)

p = ggplot(df.changes, aes(x=year,y=`Annual % Change`,col=Enforcement)) + 
  geom_abline(slope=0,intercept=0,size=1.5,col='black') + 
  geom_line(size=1) + 
  scale_y_continuous(breaks = scales::pretty_breaks(15),labels = percent) +
  theme_fivethirtyeight() + 
  scale_colour_tableau() + 
  ggtitle("Annual % Change in Incidents")
p
```

```{r}
df.changes = data %>% arrange(year)
df.changes$year = as.character(df.changes$year)

df.changes = df.changes %>%
  mutate_if(is.numeric, pct.change) %>%
  na.omit() %>%
  mutate_if(is.numeric,cumsum) %>%
  gather(Enforcement,`Annual % Change`,apprehended:returns)

df.changes$year = as.integer(df.changes$year)
#df.changes$decade = (df.changes$year %/% 10) * 10
df.changes$Enforcement = factor(df.changes$Enforcement)

p = ggplot(df.changes, aes(x=year,y=`Annual % Change`,col=Enforcement)) + 
  geom_abline(slope=0,intercept=0,size=1.5,col='black') + 
  geom_line(size=1) + 
  scale_y_continuous(breaks = scales::pretty_breaks(15), labels = percent) +
  theme_fivethirtyeight() + 
  scale_colour_tableau() + 
  ggtitle("Cumulative % Change in Incidents")
p

```



```{r}
df.clean = df
df.clean[is.na(df.clean)] = 0
df.clean = df.clean[,colSums(df.clean) >= 20]
df.total_only = df.clean[, grepl("total|year|removals|returns|apprehended", names(df.clean))]
df.no_total = df.clean[, !grepl("total", names(df.clean))]
```

```{r}
cormat <- round(cor(df.total_only %>% select(-year)),2)
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```


```{r}
cormat <- round(cor(df.no_total %>% select(-year)),2)
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```


```{r, fig.height=8}
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

# Reorder the correlation matrix
cormat <- round(cor(df.no_total %>% select(-year)),2)
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 20, hjust = 1),
    axis.text.y = element_text(size=20))+
 coord_fixed()
# Print the heatmap
print(ggheatmap)


```


```{r}
library(ggpmisc)
formula = y ~ x
p = ggplot(df)
p1 = p + geom_line(aes(x=year,y=racialbias_total)) + theme_economist()
p2 = p + geom_line(aes(x=year,y=apprehended)) + theme_economist()
p3 = p + geom_point(aes(x=racialbias_total, y=apprehended)) + theme_economist() + 
  geom_smooth(aes(x=racialbias_total, y=apprehended), method = lm, se=FALSE) + 
  stat_poly_eq(aes(x=racialbias_total, y=apprehended,label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
             label.x.npc = "right", label.y.npc = 0.15,
             formula = formula, parse = TRUE, size = 3)
grid.arrange(p1,p2,p3,ncol=1)
```




```{r}
formula = y ~ x
p = ggplot(df)
p1 = p + geom_line(aes(x=year,y=anti_protestant)) + theme_economist()
p2 = p + geom_line(aes(x=year,y=anti_islamic)) + theme_economist()
p3 = p + geom_point(aes(x=anti_protestant, y=anti_islamic)) + theme_economist() + 
  geom_smooth(aes(x=anti_protestant, y=anti_islamic), method = lm, se=FALSE) + 
  stat_poly_eq(aes(x=anti_protestant, y=anti_islamic,label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
             label.x.npc = "right", label.y.npc = 0.15,
             formula = formula, parse = TRUE, size = 3)
grid.arrange(p1,p2,p3,ncol=1)

```



```{r}
pct.change <- function(x) {(x-lag(x))/lag(x)}
# Change Rates
df.changes = df %>% arrange(year)
df.changes$year = as.character(df.changes$year)
df.changes = df.changes %>%
  mutate_if(is.numeric, pct.change) %>%
  filter(year != '2000')
df.changes$year = as.integer(df.changes$year)
df.changes
p = ggplot(df.changes)
p1 = p + geom_line(aes(x=year,y=racialbias_total)) + theme_economist()
p2 = p + geom_line(aes(x=year,y=apprehended)) + theme_economist()
p3 = p + geom_point(aes(x=racialbias_total, y=apprehended)) + theme_economist() + 
  geom_smooth(aes(x=racialbias_total, y=apprehended), method = lm, se=FALSE) + 
  stat_poly_eq(aes(x=racialbias_total, y=apprehended,label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
             label.x.npc = "right", label.y.npc = 0.80,
             formula = formula, parse = TRUE, size = 3)
grid.arrange(p1,p2,p3,ncol=1)
```

```{r}
pct.change <- function(x) {(x-lag(x))/x}
# Change Rates
df.changes = df %>% arrange(year)
df.changes$year = as.character(df.changes$year)
df.changes = df.changes %>%
  mutate_if(is.numeric, pct.change) %>%
  filter(year != '2000')
df.changes$year = as.integer(df.changes$year)
df.changes

df.changes$year = as.character(df.changes$year)
df.change.total = df.changes %>%
  mutate_if(is.numeric, cumsum)
df.change.total$year = as.integer(df.change.total$year)
df.change.total

p = ggplot(df.change.total)
p1 = p + geom_line(aes(x=year,y=racialbias_total)) + theme_economist()
p2 = p + geom_line(aes(x=year,y=apprehended)) + theme_economist()
p3 = p + geom_point(aes(x=racialbias_total, y=apprehended)) + theme_economist() + 
  geom_smooth(aes(x=racialbias_total, y=apprehended), method = lm, se=FALSE) + 
  stat_poly_eq(aes(x=racialbias_total, y=apprehended,label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
             label.x.npc = "right", label.y.npc = 0.15,
             formula = formula, parse = TRUE, size = 3)
grid.arrange(p1,p2,p3,ncol=1)
```

