
```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#knitr::opts_chunk$set(echo = TRUE)
```

### Trend of US Immigration Enforcement

In the coming month I'll be digging into the immigration enforcement data posted on [data.world](https://data.world/stoltzmaniac/insights-from-the-usa). I encourage anyone to take this data and either add to the project or to do something on their own. I will be bringing in external data sources to merge as well (which I did for this first plot).

Anyway, if you're only here for a "high-level nugget" of information, the basic thing you can see is:  

#### Things have changed since 1925!  

Here's a look at the immigration enforcement data with the party of the POTUS shown in red and blue:  

----

```{r execute script}
library(data.world)
library(reshape2)
library(tidyverse)
library(gridExtra)
library(scales)
library(GGally)
library(lubridate)
library(ggthemes)

# Reset API token needs to be changed once in a while
#data.world::set_config(save_config(auth_token = "YOUR API TOKEN"))

# Datasets are referenced by their URL or path
dataset_key <- "https://data.world/stoltzmaniac/insights-from-the-usa"

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", "immigrationenforcement"))
data = data.world::query(qry, dataset = dataset_key)

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", '02_2_fbi_hate_crime_statistics_comparison_2000_2015_all'))
hate_crimes = data.world::query(qry, dataset = dataset_key)

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", 'us_population_1990_to_2016'))
us_population = data.world::query(qry, dataset = dataset_key)

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", 'gdp'))
us_gdp = data.world::query(qry, dataset = dataset_key)


df = us_gdp %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  mutate(maxGDP = max(gdp)) %>%
  summarise(gdp = mean(maxGDP)) %>%
  left_join(us_population, by='year') %>%
  mutate(gdp.per.capita = round(1e9*gdp / population,0))


df.population.enforcement = data %>%
  left_join(us_population, by = 'year') %>% 
  filter(year >= 1927) %>%
  rename(us_population = population) %>%
  arrange(year)

df = hate_crimes %>%
  left_join(data, by = 'year') %>% 
  arrange(year)

```


```{r}

```




```{r immigration enforcement plots, fig.height=10, fig.align='center'}
pct.change <- function(x) {(x-lag(x))/lag(x)}
df.arranged = df.population.enforcement %>% arrange(year)
df.arranged$year = as.character(df.arranged$year)

df.changes = df.arranged %>%
  mutate_if(is.numeric, pct.change) %>%
  na.omit() %>% 
  gather(Metric,`Annual % Change`,apprehended:us_population) %>%
  arrange(year)

df.changes$year = as.integer(df.changes$year)
df.changes$Metric = factor(df.changes$Metric)

p1 = ggplot(df.changes, aes(x=year,y=`Annual % Change`,col=Metric)) + 
  geom_abline(slope=0,intercept=0,size=1.5,col='black') + 
  geom_line(size=1) + 
  scale_y_continuous(breaks = scales::pretty_breaks(15),labels = percent) +
  theme_fivethirtyeight() + 
  scale_colour_tableau() + 
  ggtitle("Annual % Change")


df.arranged = df.population.enforcement %>% arrange(year)
df.arranged$year = as.character(df.arranged$year)

df.changes = df.arranged %>%
  na.omit() %>% 
  gather(Metric,Values,apprehended:us_population) %>%
  select(Metric,year,Values) %>%
  arrange(year) %>%
  group_by(Metric) %>%
  mutate(difference_from_first = Values - first(Values), 
         percent_diff_from_first = round(difference_from_first /first(Values),2))


df.changes$year = as.integer(df.changes$year)
p2 = ggplot(df.changes, aes(x=year,y=percent_diff_from_first,col=Metric)) + 
  geom_line(size=1) + 
  scale_y_continuous(breaks = scales::pretty_breaks(15), labels = percent) +
  theme_fivethirtyeight() + 
  scale_colour_tableau() + 
  ggtitle("Since 1925 % Change")

grid.arrange(p1,p2,ncol=1)
```




```{r}
enforcement.pop.adjusted = df.population.enforcement %>%
  mutate(apprehended = (apprehended / us_population)*1000,
         removals = (removals / us_population)*1000,
         returns = (returns / us_population)*1000) %>%
  select(year,apprehended,removals,returns) %>%
  gather(Enforcement, Incidents.Per.1K.Population, apprehended:returns)

p = ggplot(enforcement.pop.adjusted,aes(x=year,y=Incidents.Per.1K.Population,col=Enforcement))
p + geom_line()
```

