
```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#knitr::opts_chunk$set(echo = TRUE)
```

### Trend of US Immigration Enforcement

In the coming month I'll be digging into the immigration enforcement data posted on [data.world](https://data.world/stoltzmaniac/insights-from-the-usa). I encourage anyone to take this data and either add to the project or to do something on their own. I will be bringing in external data sources to merge as well (which I did for this first plot).

Anyway, if you're only here for a "high-level nugget" of information, the basic thing you can see is:  

#### Things have changed since 1925!  

Here's a look at the immigration enforcement data with the party of the POTUS shown in red and blue:  

----

```{r execute script}
library(data.world)
library(reshape2)
library(tidyverse)
library(gridExtra)
library(scales)
library(GGally)
library(lubridate)
library(ggthemes)

# Reset API token needs to be changed once in a while
#data.world::set_config(save_config(auth_token = "YOUR API TOKEN"))

# Datasets are referenced by their URL or path
dataset_key <- "https://data.world/stoltzmaniac/insights-from-the-usa"

#qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", "immigrationenforcement"))
#data = data.world::query(qry, dataset = dataset_key)

#qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", '02_2_fbi_hate_crime_statistics_comparison_2000_2015_all'))
#hate_crimes = data.world::query(qry, dataset = dataset_key)

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", 'us_population_1990_to_2016'))
us_population = data.world::query(qry, dataset = dataset_key)

qry = data.world::qry_sql(sprintf("SELECT * FROM `%s`", 'gdp'))
us_gdp = data.world::query(qry, dataset = dataset_key)


df = us_gdp %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  mutate(maxGDP = max(gdp)) %>%
  summarise(gdp = mean(maxGDP)) %>%
  left_join(us_population, by='year') %>%
  mutate(gdp.per.capita = round(1e9*gdp / population,0))

```


```{r, fig.height=6}
p = ggplot(df,aes(x=year)) + theme_economist()

p1 = p + geom_line(aes(y=gdp),size=1.5) + 
  ggtitle("US GDP in Billions ($)") + 
  scale_y_continuous(labels=dollar) +
  scale_x_continuous(breaks = seq(min(df$year)-1,max(df$year),by=10)) + 
  ylab('')

p2 = p + geom_line(aes(y=population),size=1.5) + 
  ggtitle("US Population") + 
  scale_y_continuous(labels=comma) + 
  scale_x_continuous(breaks = seq(min(df$year)-1,max(df$year),by=10)) + 
  ylab('')
p3 = p + geom_line(aes(y=gdp.per.capita),size=1.5) + 
  ggtitle("US GDP Per Capita") +
  scale_y_continuous(labels=dollar) + 
  scale_x_continuous(breaks = seq(min(df$year)-1,max(df$year),by=10)) + 
  ylab('')

grid.arrange(p1,p2,p3,ncol=1)
```

